MLH1 Results Outline

### 2020-02-12 doc/Results.rmd
end o day o

load(file="~./MLH1repo/data/CleanBivData_3.2.20.RData")

@@

@@

@@

# gwRR estimates

make table of the strain averages

@@

#variance plots
musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)

Mol.female.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)

Dom.female.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)

@@

@@

@@

@@

sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM")
mean.all.other <- mean(sans_high_female$mean_co)

#how much greater is MSM female
MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
mean.MSN.female <- mean(MSM_female$mean_co)

sans_MSM_female <- female_mouse.avs_for.analysis %>% filter(strain != "MSM")
mean.all.other <- mean(sans_MSM_female$mean_co)

mean.MSN.female / mean.all.other

@@

The range of strain means is `r min()`

the ranges of mouse means for each strain is X

MSM, G, LEW have significant strain effects.

@@

Follow up on this finding -- included more strains which were excluded from the HetC analysis.

> Is there a strain effect in Dom?

> How different are the Musc strains?

test a model for fixed effects of strains within musc. The purpose is to distinguish musc strains from low 'Dom' level strains.

@@

(Try removing the alpha legend)

This is the table and plot of new Musc strains

`r kable(table( droplevels( Musc.male_mouse.av$strain)  ))`

Given the plot, make note of the potential outlier mice.  The strains have unequal variance

KAZ: 10jul19_KAZ_m3, 10jul19_KAZ_m1, 10jul19_KAZ_m2

PWD: 24sep14_PWD_m1, 18may15_PWD_m1?

@@

@@

#1. determine which strain effects are sig
#M2
lm.male.strain <- glm(mean_co ~ strain, data= just.male_mouse.av)
lm.male.strain_sum<- summary(lm.male.strain)

@@

SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
#mean(SKIVE_male$mean_co) /  
  mean(sans_LewRec_male$mean_co)

@@

the ranges of means are X

PWD, MSM and SKIVE have strain significant strain effects so they will be designedated into the high rec group.

@@

In order to understand the variance within an evolutionary framework we fit the mouse gwRRs to a mixed model, which specified subspecies, sex and the interaction as fixed effects. The subspecies effect - is a proxy for measuring the degree of divergence across subspecies. The strains were coded as a random effect to -simulate / approximate / the standing genetic variation across subspecies.

@@

#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
```

@@

`r kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`

1. $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$

@@

MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
#summary(MLH1.MM.M1)

MM.Fixed.pval <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value

@@

2. $$mouse \ av \ meteric ~=~ subsp * sex * strain + \varepsilon$$

@@

summ <- summary(glm.strain.fixed.full)

@@

Strain G is significant (p=0.0005),

@@

3. $$mouse \ av \ meteric ~=~ sex * strain + \varepsilon$$

Used after it's apparent strain has a significant effect - another post-hoc examination for strain as a fixed effect.

$$mouse \ av \ metric ~=~ subsp * sex * strain + \varepsilon$$

@@

#below table is broken
as.data.frame( round( summary(strain.fixed)$coefficients, 6))

@@

## Within Mouse Variance in CO Count per Cell

@@

**todo**

Test if dissection date is more significant predictor than strain (for females) (Effect doc)
Also test how the number of 0CO bivalents per cell affects mouse with variance. (lm)

@@

> Modeling Within Mouse Variance

(use the same models are mean_co, )

XXX (old 1.)	HetC.MixedModel.HQ <- lme(mouse.metric ~ subsp*sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )

1.(new 1)	Reduced.strain.HQ <- lmer(mouse.metric ~ subsp * sex+ (1|strain), data=DF.HetC.MixedModel.HQ)

3.	lm(mouse.metric ~ subsp \* sex \* strain, data=DF.HetC.MixedModel.HQ)

4.	lm(mouse.metric ~ sex * strain, data=DF.HetC.MixedModel.HQ)

@@

```{r M1.MLH1.var_cv, echo=FALSE, include=FALSE}
#modeling within mouse variance
#use  DF.HetC.MixedModel.HQ

#HQ var.CO
MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
#summary(MLH1var_M1)

drop1(MLH1var_M1, test="Chisq") #interaction term NS
#when switched to additive fixed, sex is very sig

anova(MLH1var_M1)#sex explains most variance...

exactRLRT(MLH1var_M1) #strain
as.data.frame( round(summary(MLH1var_M1)$coefficients, 6))

#CV
MLH1cv_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
#summary(MLH1cv_M1)

drop1(MLH1cv_M1, test="Chisq") #interaction term NS
#when switched to additive fixed, sex is very sig
anova(MLH1cv_M1)  #sex F value is large
exactRLRT(MLH1cv_M1) #strain

as.data.frame( round(summary(MLH1cv_M1)$coefficients, 6))

#USE Q12 (Q123) table
#Q12_mouse_table_w.Ages
#Q12 var
MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )

drop1(MLH1var_M1_Q12, test="Chisq")#sig
anova(MLH1var_M1_Q12)
exactRLRT(MLH1var_M1_Q12)#NS

as.data.frame( round(summary(MLH1var_M1_Q12)$coefficients, 6))

#Q12 cv
MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )

drop1(MLH1cv_M1_Q12, test="Chisq")#NS
anova(MLH1cv_M1_Q12)#sex large F
exactRLRT(MLH1cv_M1_Q12)#NS

as.data.frame( round(summary(MLH1cv_M1_Q12)$coefficients, 6))

The fixed sex effect has the largest effect sizes. All are negative (for male).  (I'm not sure if I should switch to testing the additive fixed effects by themselves) - I'm pretty sure the outcome would be that the sex effect gets very large in those models.
The strain effects are only slightly significant for within mouse variance measures in the full data set (includes low quality cells)

$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$

```{r M2.MLH1.var, echo=FALSE, include=FALSE}

@@

For LEW - females must have high variance,  for CV Lew strain alone has increased cv. Consider plotting the variance - and diving deeper into this pattern
Most of the strain specific differences in within mouse variance go away in the Q12 data set.

@@

```{r M3.MLH1.var, echo=FALSE, include=FALSE}

@@

#summary(MLH1var_M4)

drop1(MLH1var_M4, test="Chisq")
anova(MLH1var_M4)
as.data.frame( round(summary(MLH1var_M4)$coefficients, 6))

@@

#summary(MLH1cv_M4)

anova(MLH1cv_M4)  
as.data.frame( round(summary(MLH1cv_M4)$coefficients, 6))

### Q12 var
MLH1var_M4_Q12 <- lm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )

#summary(MLH1var_M4_Q12)

anova(MLH1var_M4_Q12)
as.data.frame( round(summary(MLH1var_M4_Q12)$coefficients, 6))

@@

anova(MLH1cv_M4_Q12)

as.data.frame( round(summary(MLH1cv_M4_Q12)$coefficients, 6))

#### Summary witin mouse Variance

The general pattern across all the models, is that sex is a significant effect for all models. The sex-subsp interaction effect p values were larger and only significant for one of the models.

@@

<!--
For later analysis.  Not sure what these are good for.

<!--  come back to Fig1
# Figure 1

work more on figure 1. Main figure, MLH1 strain avs on top, chrm proportion bar plot on bottom

strain.table.HQ_Fig <- ddply(MLH1_data, c("strain", "sex"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  =length(nMLH1.foci),
                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                       #Standard error (mice vs cells)
                        se  = round(sd / sqrt(Ncells), 3)
    )

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male

   #segments
  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #cast
  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
  #outgroups
  annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
  #label
  annotate("text", x = c(3,8.5,10,12, 14), #list of all x points 
           y = c(18.5,18.5,33,18.5, 18.5),

   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
           fontface = 'italic', hjust = 0.5, size=3.5)

What are the steps the each chunks walk through.

>>>>

load(file="~./MLH1repo/data/CleanBivData_11.2.20.RData")

@@

# Main Figure

strain.table.HQ_Fig.og <- ddply(MLH1_data, c("strain", "sex"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  =length(nMLH1.foci),
                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                       #Standard error (mice vs cells)
                        se  = round(sd / sqrt(Ncells), 3)
    )

strain.table.HQ_Fig <-  strain.table.HQ_Fig.og %>% filter(strain != "F1" & strain != "PERC" &
                       strain != "TOM" &  strain != "AST" & strain != "CZECH" & strain != "CAROLI" & strain != "HMI")

#adjust order of plot here
strain.table.HQ_Fig$strain <- factor(strain.table.HQ_Fig$strain,levels =c(
     "MSM", "MOLF",   
  "PWD", "SKIVE",  "KAZ",
  "WSB","G","LEW","PERC",
                     "CAST",
                         "SPRET","SPIC" ), order=T )

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
#remove y labfor male

   #segments
#  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
#  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
#  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #cast
#  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
  #outgroups
 # annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
  #label

 #   annotate("text", x = c(2,4.4,6,8, 10), #list of all x points 
   #        y = c(18.5,18.5,33,18.5, 18.5),

  # label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
    #       fontface = 'italic', hjust = 0.5, size=3.5)

This is a draft of the main figure. Outgroups are not included in the analysis below.

@@

```{r maw.fig1s, echo=FALSE, include=FALSE}

strain.table.HQ_Fig_F <-  strain.table.HQ_Fig %>% filter(sex == "female") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" )

Fig1_Female <- ggplot(strain.table.HQ_Fig_F, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point( size=3, position=position_dodge(.5))+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())

#strain.table.HQ_Fig.og

strain.table.HQ_Fig_M <-  strain.table.HQ_Fig.og %>% filter(sex == "male") %>% filter(strain != "CAST" & strain != "SPRET" & strain != "SPIC" & strain != "PERC" & strain != "F1" & strain != "HMI" & strain != "CAROLI")

strain.table.HQ_Fig_M$strain <- factor(strain.table.HQ_Fig_M$strain,levels =c(
     "MSM", "MOLF",   
  "PWD", "SKIVE",  "KAZ", "TOM", "AST","CZECH",
  "WSB","G","LEW"), order=T )

Fig1_Male <- ggplot( strain.table.HQ_Fig_M, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point(size=3, position=position_dodge(.5))+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())

@@

# Genome wide recombiantion rate estimates

@@

@@

G and LEW have a higher strain means compared to WSB. MSM (and PWD) also has an elevated mean.

The variance across mouse means for each strain is also variable across strains.

In the dom  subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in WSB, `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in G, `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in LEW.

In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in PWD, `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in KAZ.

The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in MSM.

Two glm models were run to test the effects of strain and subspecies.

@@

round(summry.female.strain.effects$coefficients[,c(1,4)],5)

@@

round(lm.female.strain_sum$coefficients[,c(1,4)],5)

@@

LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
mean.LEW.female <- mean(LEW_female$mean_co)

#how much greater is MSM female
MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
mean.MSM.female <- mean(MSM_female$mean_co)

sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
mean.all.other <- mean(sans_high_female$mean_co)

#sans_MSM_female / mean.all.other

@@

Above are the coefficents for the two glm's of the female specific mouse MLH1 averages, which include subspecies and strain as fixed effects. G has the most consistant significant strain effects in both models. MSM has a pretty low p value in the second model, LEW has a slightly significant pvalue.

G is `r mean.G.female / mean.all.other` higher the the other means. LEW is `r mean.LEW.female / mean.all.other` higher  and MSM is `r mean.MSM.female / mean.all.other`. These three will be designated as 'moderate high rec' strains for later sex specific analysis.

@@

The male specific analysis was done to assess the variance across strains. The plots below illustrate mouse level means for MLH1 per cell serperated by subspecies.

@@

- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )`

- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` and Molossinus.

- While there is alot of variance within strains, a general pattern is that PWD, SKIVE and MSM can be classified as 'high rec' strains, there strain averages are

@@

Two models were fitted to the male specific data to test the effects of strain and subspecies.

@@

#1. determine which strain effects are sig

#M2
lm.male.strain <- glm(mean_co ~ strain, data= just.male_mouse.av)
lm.male.strain_sum<- summary(lm.male.strain)

round(lm.male.strain_sum$coefficients[,c(1,4)], 5)

@@

SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
#mean(SKIVE_male$mean_co) /

@@

The strain average for PWD, MSM and SKIVE are  ` r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively. Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.

@@

In order to understand the variance within an evolutionary framework we fit the mouse gwRRs to a mixed model, which specified subspecies, sex and the interaction as fixed effects. The subspecies effect - is a proxy for measuring the degree of divergence across subspecies. The strains were coded as a random effect to approximate standing genetic variation across subspecies ranges.

@@

#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))

all.cell.MM.HQ <- ddply(.data=DF.HetC.MixedModel.HQ,
                 .(sex, strain),
                 summarize,
                 all.cells = sum(Ncells)
)

@@

`r kable(table( DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))`

add number of cells

`r kable(table( all.cell.MM.HQ$sex, all.cell.MM.HQ$strain))`

 $$mouse \ av \ meteric ~=~ subsp * sex  + rand(1|strain) + \varepsilon$$

@@

MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

#summary(MLH1.MM.M1)

MM.Fixed.pval <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]

anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)
anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)

MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value

@@

 $$mouse \ av \ meteric ~=~ subsp * sex * strain + \varepsilon$$

@@

summ <- summary(glm.strain.fixed.full)
#table of just coeff anf pvalue
round(summ$coefficients[,c(1,4)],5 )

@@

Strain G is significant (p=0.0005), (remember female result)

@@

@@

summ.strain.fixed <- summary(strain.fixed)

round(summ.strain.fixed$coefficients[,c(1,4)],5)

#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)

#as.data.frame( round( summary(strain.fixed)$coefficients, 6))

@@

# Within Mouse Variance in CO Count per Cell

@@

@@

```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5}

#variance plots
musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
musc.female.VAR.plot

Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Mol.female.VAR.plot

Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Dom.female.VAR.plot

Check that the one PWD female with like 0 variance... wtf, that mouse 8oct14_PWD_f2 has 3 cells. make sure it's removed from this table.

```{r mouse.var.plots.male, echo=FALSE, fig.width = 4, fig.height=2.5}

#variance plots
musc.male.VAR.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Musc male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
musc.male.VAR.plot

Mol.male.VAR.plot <- ggplot(Mol.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Mol male of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Mol.male.VAR.plot

Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Dom.male.VAR.plot

from the Mixed models, there should be 4, 1-2 models from the full dataset for var and cv, and 3-4 from a higher quality data set for var and cv. Only a subset of the model outputs will be shown.

For the mixed models the indiviual fixed effects are tested by a anova of the full and reduced models. The interaction fixed effect pvalue comes from drop1() a LTR test.

@@

```{r M1.MLH1.var_cv, echo=FALSE}
#modeling within mouse variance
#use  DF.HetC.MixedModel.HQ

#HQ var.CO
MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

#summary(MLH1var_M1)
#drop1(MLH1var_M1, test="Chisq") #interaction term NS
#when switched to additive fixed, sex is very sig

#anova(MLH1var_M1, MLH1var_M1.reduce.sex)
#anova(MLH1var_M1, MLH1var_M1.reduce.subp)

#exactRLRT(MLH1var_M1) #strain
#as.data.frame( round(summary(MLH1var_M1)$coefficients, 6))

#CV
MLH1cV_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1cV_M1.reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1cV_M1.reduce.subp <- lmer( cV ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

#drop1(MLH1cV_M1, test="Chisq") #interaction term NS #when switched to additive fixed, sex is very sig
#anova(MLH1cV_M1, MLH1cV_M1.reduce.sex)
#anova(MLH1cV_M1, MLH1cV_M1.reduce.subp)
#exactRLRT(MLH1cV_M1) #strain

#USE Q12 (Q123) table
#Q12_mouse_table_w.Ages
#Q12 var
MLH1var_M1_Q12 <- lmer( var ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1var_M1_Q12reduce.sex <- lmer( var ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1var_M1_Q12reduce.subp <- lmer( var ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )

#drop1(MLH1var_M1_Q12, test="Chisq")#sig

#anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.sex)
#anova(MLH1var_M1_Q12, MLH1var_M1_Q12reduce.subp)
#exactRLRT(MLH1var_M1_Q12)#NS

#Q12 cv
MLH1cv_M1_Q12 <- lmer( cV ~ subsp * sex +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1cv_M1_Q12reduce.sex <- lmer( cV ~ subsp  +(1|strain), data=Q12_mouse_table_w.Ages )
MLH1cv_M1_Q12reduce.subp <- lmer( cV ~ sex  +(1|strain), data=Q12_mouse_table_w.Ages )

#drop1(MLH1cv_M1_Q12, test="Chisq")#NS

pval.sig <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.sex)
CV.anova.p <- pval.sig$`Pr(>Chisq)`[2]

# <- anova(MLH1cv_M1_Q12, MLH1cv_M1_Q12reduce.subp)

#exactRLRT(MLH1cv_M1_Q12)#NS

#make a table of the Mixed model pvalues, which are caluclated for subp, sex, sex:subsp and strain across 4 models ()
```

**General pattern is that sex has a very significant effect.** ( on the order of `r round(CV.anova.p, 6) ` )  this applies to different measures of within animal (across cell) variance and also to subsetted data with higher quality cells.

```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
#not including this model for now / brivity
#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$

@@

<!-- insert the sex specific variance effect here / maybe also the plots of variance -->

For LEW - females must have high variance,  for CV Lew strain alone has increased cv. Consider plotting the variance - and diving deeper into this pattern
Most of the strain specific differences in within mouse variance go away in the Q12 data set.  (These strain effects could indicate that the distribution of across cell variance in CO number evolves.)

The second set of models analyzed have this form.

@@

```{r M3.MLH1.var, echo=FALSE}

@@

#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)
#as.data.frame( round(summary(MLH1var_M4)$coefficients[,c(1,4)], 6))

@@

#summary(MLH1cv_M4)
#as.data.frame( round(summary(MLH1cv_M4)$coefficients, 6))

### Q12 var
MLH1var_M4_Q12 <- lm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
#summary(MLH1var_M4_Q12)
#as.data.frame( round(summary(MLH1var_M4_Q12)$coefficients, 6))
#as.data.frame( round(summary(MLH1var_M4_Q12)$coefficients[,c(1,4)], 6))

@@

#anova(MLH1cv_M4_Q12)

#as.data.frame( round(summary(MLH1cv_M4_Q12)$coefficients, 6))

The glm's support the mixed models with the sex effect being the consistant significant effect

#### Summary for witin mouse Variance

The general pattern across all the models, is that sex is a significant effect for all models. **General pattern is that sex has a very significant effect.** The sex-subsp interaction effect p values were larger and only significant for one of the models.

The models run over higher quality cells, suggest that staining or technical error in female data is unlikely to be the only cause for increased variance of number of MLH1 foci per oocyte.

@@

#make DMC1 table
DMC1.mean.obs.L <- ddply(.data=DMC1.DF[DMC1.DF$stage == "L",],
                 .(category),
                 summarize,

                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.L$stage <- "L"

DMC1.mean.obs.Z <- ddply(.data=DMC1.DF[DMC1.DF$stage == "Z",],
                 .(category),
                 summarize, 
                 ncells = length(fileName),
                 mean.count = mean( total.count)
)
DMC1.mean.obs.Z$stage <- "Z"

DMC1.mean.table <- rbind(DMC1.mean.obs.L,DMC1.mean.obs.Z)

#write.table(DMC1.mean.table, "~./MLH1repo/DMC1.mean.table.csv", sep=",",
#            row.names = FALSE)

DMC1.mean.table

DMC1.DF$rec_group <- ifelse(grepl("WSB", DMC1.DF$strain), "Low",
                          ifelse(grepl("G",DMC1.DF$strain), "Low",

              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
            ifelse(grepl("PWD", DMC1.DF$strain), "High",
               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))

DMC1.histogram
#checking equal variance, https://www.statmethods.net/stats/anovaAssumptions.html
#bartlet's test, Levine's test

#The counts look close to normal.
#Potentially figure 2. Using a function to #display the p-vales for mean comparison

#change order

DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("LEW",
                      "WSB","G", "KAZ", "SKIVE","PWD","MSM"), order=T )

DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
  geom_boxplot(aes(alpha=.5))+geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

DMC1.boxplot
```

```{r plot.stage.ttest, echo=FALSE, warning=FALSE}

anno_df_stage = compare_means(total.count ~ stage, group.by = "rec_group", data = DMC1.DF[DMC1.DF$stage != "P",],method = "t.test") %>% mutate(y_pos = 300)
#

DMC1.bxlt_Z_to_L <- ggplot(DMC1.DF[DMC1.DF$stage != "P",], aes(x=stage, y=total.count)) + 
  geom_boxplot(position=position_dodge()) + 
  geom_point(aes(color=rec_group), position=position_jitterdodge()) +

  #facet_wrap(~stage, nrow=1) +

  ggsignif::geom_signif(
    data=anno_df_stage, 
    aes(xmin=group1, xmax=group2, annotations=p.adj, y_position=y_pos), 
    manual=TRUE
  )+ggtitle("Comparison of DMC1 means across high and low recombing groups")+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

DMC1.bxlt_Z_to_L

#Musc.mouse.av <-  Mouse.Table_BivData_4MM %>%  filter(subsp == "Musc")

ttest.ZtoL <- t.test(DMC1.DF$total.count[(DMC1.DF$stage == "L")],
               DMC1.DF$total.count[(DMC1.DF$stage == "Z")]  )

DMC1.High.goup <-  DMC1.DF %>%  filter(rec_group == "High")

ttest.ZtoL.high <- t.test(DMC1.High.goup$total.count[(DMC1.High.goup$stage == "L")],
               DMC1.High.goup$total.count[(DMC1.High.goup$stage == "Z")]  )

DMC1.Low.goup <-  DMC1.DF %>%  filter(rec_group == "Low")

ttest.ZtoL.Low <- t.test(DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "L")],
               DMC1.Low.goup$total.count[(DMC1.Low.goup$stage == "Z")]  )
```

the pvalues for the differences between time points are  `r ttest.ZtoL$p.value`,  `r ttest.ZtoL.high$p.value` , and  `r ttest.ZtoL.Low$p.value` for all observations, the high rec group and the low group respectively.

```{r plot.t.test, echo=FALSE}

  #this will add neat bars, but still need to figure out what groups ect it's actually using
#THE stat_compare function adds a arm thing and pvalue for 
#  stat_compare_means(comparisons = my_comparisons)

              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
            ifelse(grepl("PWD", DMC1.DF$strain), "High",
               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))

DMC1.DF$rec_group <- as.factor(DMC1.DF$rec_group)

# annotation table with adjusted pvals and y-position of the labels
